hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2024-12-09 inst/manuscript/paper.Rmd
copy edits

# take temps from a random normal dist with mean 10 sd 3 
# for every date time combo in dates and append to column (temp) in year_sim
year_sim$temp <- rnorm(n = length(dates), mean = 10, sd = 3) %>%
  abs()

@@

Below, we demonstrate how the `fit_model()` function may be used to create custom parameterizations for species beyond the Salmonids in the `model_table` included in the package. We include parameterizations from three warm-water species to demonstrate the `fit_model()` utility for fishes beyond the scope of the original effective value approach. We include parameterizations for commonly cultured sportfishes including Smallmouth Bass (*Micropterus dolomieu*) from @webster1948, Channel Catfish (*Ictalurus punctatus*) from @small2001, and Lake Sturgeon (*Acipenser fulvescens*) from @smith2005.

@@

Note the model the *R^2^* fit from the models below. You can see the generally all preform well and are in line with values from model 2 of @beacham1990.

@@

```{r}
sockeye_hatch_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  dev.type = "hatch"
  )

sockeye_emerge_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  dev.type = "emerge"
  )

@@

`WI_hatch$days2done` outputs the predicted days to hatch or emerge.

`WI_hatch$dev.period` is a 1x2 dataframe with the dates corresponding to when your fish's parent spawned (which you input with `predict_phenology(spawn.date = ...)`) and the date when the fish is predicted to hatch or emerge.

>>>>

# take temps from a random normal dist with mean 10 sd 3 
# for every date time combo in dates and append to column (temp) 
# in year_sim
year_sim$temp <- rnorm(n = length(dates), mean = 10, sd = 3) %>%
  abs()

@@

Below, we demonstrate how the `fit_model()` function may be used to create custom parameterizations for species beyond the Salmonids in the `model_table` included in the package. We provide parameterizations for three warm-water species to demonstrate the `fit_model()` utility for fishes beyond the scope of the original effective value approach. The parameterizations are for commonly cultured sport- and commercial fishes including Smallmouth Bass (*Micropterus dolomieu*) from @webster1948, Channel Catfish (*Ictalurus punctatus*) from @small2001, and Lake Sturgeon (*Acipenser fulvescens*) from @smith2005.

@@

Note the *R^2^* fit from the models below. You can see they generally all preform well and are in line with values from model 2 of @beacham1990.

@@

```{r}
sockeye_hatch_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  dev.type = "hatch"
  )
```

@@

`WI_hatch$days2done` outputs the predicted days to develop.

`WI_hatch$dev.period` is a 1x2 dataframe with the dates corresponding to when your fish's parent spawned (which you input with `predict_phenology(spawn.date = ...)`) and the date when the fish is predicted todevelop.
